The Concurrent Calculi Formalisation Benchmark

### 2023-10-12 paper/paper.tex
Update references.bib

As with the challenges that came before, this one is intended to be
useful on its own, but not to exhaustively cover every possible
problem in the area. The current challenge can be extended in two
dimensions: the first is taking the existing challenges further.
And the second is devising new challenges not covered by the current
benchmarks.

The proposed challenges try to minimise the overlap of the techniques
they exercise, however, most systems will combine some or all aspects
of the challenges. An interesting second round of formalisations could
entail combining the features of the challenges. In the following
table we consider the following three additions: First the addition of
choice to the type level and the process level. Second, the addition
of recursion (and recursion types for typed systems). And finally, the
addition of channel delegation to those calculi that do not already
support it.

The following table shows which extension should be suitable (that we
mark with $+$) for each challenge:

@@

The other dimension to extend these calculi corresponds to systems
that are not covered by our challenges. Extending in this dimension
means proposing new challenges to cover for different features of
message passing calculi, or to cover different features of mechanised
proof.

Some interesting aspects of message passing calculi to explore in
further challenges could be: Multiparty session types, and
choreographies because their definitions and meta-theory have elements
that we do not cover here, however the elements that we do cover will
appear in these problems. Also, while we explore a form of barbed
bisimilarity one could easily design a challenge exploring different
notions of bisimilarity (barbed, or weak, etc.), and trace
equivalence.

On the side of proof assistants, the current challenge proposes
mechanising the proofs of theorems. An interesting venue to explore
would be to take advantage of other aspects of proof assistants. For
example, proof assistants are often able to produce certified code (by
using code extraction, or compiling and running their definitions).
Additionally, while having effective automation in our current
challenge is possible, a challenge could be designed with the explicit
objective of automating aspects of the proofs. And finally, a
challenge could propose the integration with other formal reasoning
tools, namely model checkers. These tools are in common use in the
field. However, combining automated proofs with proof assistants
offers the potential to ease the path for attempting larger proofs
with less effort.

Ultimately, the current challenges can be extended in several
worthwhile directions, and we look forward to a future when they
indeed are extended. Moreover, we see the current challenge as setting
the foundation for those future extensions. It is our expectation that
the solutions of those hypothetical challenges will build on solutions
to this challenge.

@@

>>>>

As with the POPLMark and POPLMark Reloaded challegnes, this challenge is
intended to be useful, but it does not aim at covering every possible
problem in the area. Our benchmark can be extended in two
dimensions: further expanding and combining the current challenges,
and addressing calculi and features not covered by the current challenges.
We discuss these dimensions in the rest if this section.

\paragraph{Extending and combining the current challenges.}
The proposed challenges try to minimise the overlap of the techniques
and aspects they exercise (linearity, scope extrusion, coinduction) --- but most
research on concurrency theory combines some or all such aspects. An interesting second round of formalisations could
entail combining the features of the challenges.

In the following table we consider the following three extensions: first, the
addition of choice to the type level and the process level. Second, the addition
of recursion (and recursion types for typed systems). And finally, the
addition of channel delegation to those calculi that do not already
support it.  We mark each suitable extension with ``$+$'':

@@

\paragraph{Addressing new calculi and new features.}
Extending in this dimension
means proposing new challenges to cover for different features of
message passing calculi, or to cover different proof mechanisation features.

Some interesting aspects of message passing calculi to explore in further
challenges could be multiparty session types, and choreographies because their
definitions and meta-theory have aspects that we do not cover here\fxerror{Such
as?} (although the aspects that we cover here do appear in these problems).
Also, one could easily design a challenge that goes beyond our coinduction
challenge, by exploring different notions of bisimilarity (barbed, or weak,
etc.), and trace equivalence.

On the side of proof assistants, the current challenge proposes
mechanising the proofs of theorems. An interesting venue to explore
would be to take advantage of other proof assistant features. For
example, proof assistants are often able to produce certified code (by
using code extraction, or compiling and running their definitions).
Additionally, while having effective automation in our current
challenge is possible, a challenge could be designed with the explicit
objective of automating aspects of the proofs. And finally, a
challenge could propose the integration with other formal reasoning
tools, namely model checkers. These tools are in common use in the
field --- and combining automated proofs with proof assistants
offers the potential to ease the path for attempting larger proofs
with less effort.

Ultimately, the current challenges can be extended in several
worthwhile directions, and we look forward to a future when they
indeed are extended. We see the current challenge as setting
the foundation for those future extensions. It is our expectation that
the solutions of those hypothetical challenges will build on solutions
to this challenge.

@@

\fxerror{TODO? Remove?}
